---
title: "Genotyping QC"
output: html_notebook
---

- See the [wiki](http://wiki.psycm.cf.ac.uk/wiki/index.php/GWAS_Quality_control_(QC)_pipeline) for QC steps

```{r setup}
library(tidyverse)
VCFindex <- read_delim("~/BTSync/FetalRNAseq/LabNotes/VCFindex.txt", 
     "\t", escape_double = FALSE, col_names = c('Sample', 'Index', 'VCF'), 
     trim_ws = TRUE)
SampleInfo <- read_delim("~/BTSync/FetalRNAseq/Github/GENEX-FB1/Data/SampleInfo.txt", 
     "\t", escape_double = FALSE, col_types = cols(Sample = col_character()), 
     trim_ws = TRUE)

key <- read_delim("/Users/heo3/BTSync/FetalRNAseq/Genome-wide\ genotyping/PLINK_040517_1054/key.txt", 
     "\t", escape_double = FALSE, col_names = c('Sample', 'IID'), 
     trim_ws = TRUE)

sample_info <- read_delim("~/BTSync/FetalRNAseq/LabNotes/sample_info.txt", 
     "\t", escape_double = FALSE, col_types = cols(BrainBankID = col_character(), 
         PCW = col_double()), trim_ws = TRUE)

sample_info <- sample_info %>% mutate(PCW=floor(PCW)) %>% 
  select(Sample=BrainBankID, Sex, PCW) %>% 
  anti_join(SampleInfo, by="Sample") %>% 
  bind_rows(mutate(SampleInfo, PCW=as.numeric(PCW))) %>% 
  full_join(key) %>% full_join(VCFindex) %>% 
  select(Sample, Index, IID, VCF, everything()) %>%
  mutate(Sex=ifelse(IID %in% c('200720170117_R05C01', '200720170117_R12C01', '200720170085_R10C02'), 'Male',
                    ifelse(IID %in% c('200720170117_R06C01', '200720170117_R11C01', '200720170117_R11C02'), 
                           'Female', Sex)))

write_tsv(sample_info, "~/BTSync/FetalRNAseq/Github/GENEX-FB2/Data/SampleInfo.txt")

```
```{bash plinkSex}
cd /Users/heo3/BTSync/FetalRNAseq/Genome-wide\ genotyping/PLINK_040517_1054

# convert to binary
~/bin/plink --file Nick_Bray_OmniExpress_230317 --make-bed --out FB_new

# remove blank samples
echo 24 > excluded_fam.txt
seq 45 48 >> excluded_fam.txt
~/bin/plink --bfile FB_new --remove-fam excluded_fam.txt --make-bed --out FB_new_excl_blank

# add Sex
grep FB_new ~/BTSync/FetalRNAseq/Github/GENEX-FB2/Data/SampleInfo.txt | cut -f 2,3,5 | perl -pe 's/Male/1/; s/Female/2/; s/NA/0/' > gender.txt  
~/bin/plink --bfile FB_new_excl_blank --update-sex gender.txt --make-bed --out FB_new_with_sex
~/bin/plink --bfile FB_new_with_sex --check-sex --out FB_new_with_sex_sexcheck
```

```{bash plinkMiss}
# Investigate missingness rates
~/bin/plink --bfile FB_new_with_sex --missing --out FB_new_with_sex_miss
```
- One sample (12972) has a call rate of 97.7%
- All others are > 99%

```{r IMISS}
IMISS = read.table("~/BTSync/FetalRNAseq/Genome-wide\ genotyping/PLINK_040517_1054/FB_new_with_sex_miss.imiss", h=T, as.is=T)
plot((1:dim(IMISS)[1])/(dim(IMISS)[1]-1), sort(1-IMISS$F_MISS), main ="Ordered individual Call Rate", xlab="Quantile", ylab="Call Rate"); grid()
filter(IMISS, F_MISS > .01) %>% left_join(sample_info)
```
- 92.3% of SNPs are complete
- 5.5% have one missing call
- 2.2% have > 1 missing call
- If I filter on missingness of 2% or less, it will remove all snps with missing calls because there are only 43 samples
```{r LMISS}
LMISS = read.table("~/BTSync/FetalRNAseq/Genome-wide\ genotyping/PLINK_040517_1054/FB_new_with_sex_miss.lmiss", h=T, as.is=T)
plot((1:dim(LMISS)[1])/(dim(LMISS)[1]-1), sort(1-LMISS$F_MISS), main ="Ordered SNP Coverage", xlab="Quantile", ylab="Coverage"); grid()

LMISS %>%
  arrange(desc(F_MISS)) %>%
  mutate(rank=row_number()) %>% 
  ggplot(aes(x=rank, y=1-F_MISS)) + geom_point()

group_by(LMISS, N_MISS) %>% summarise(n()/nrow(LMISS))

```


# Minor Allele Frequency
```{bash plinkMAF}
cd /Users/heo3/BTSync/FetalRNAseq/Genome-wide\ genotyping/PLINK_040517_1054
~/bin/plink --bfile FB_new_with_sex --freq --out FB_new_with_sex_freq
```
```{r MAF}
FREQ = read.table("~/BTSync/FetalRNAseq/Genome-wide\ genotyping/PLINK_040517_1054/FB_new_with_sex_freq.frq", h=T, as.is=T)

ggplot(FREQ, aes(x=MAF)) + geom_freqpoly(bins=44)

```

# Heterozygosity

- 6 samples have low F and high H (suggests possible sample contamination?)
```{bash plink_heterozygosity}
cd /Users/heo3/BTSync/FetalRNAseq/Genome-wide\ genotyping/PLINK_040517_1054
~/bin/plink --bfile FB_new_with_sex --het --out FB_new_with_sex_het
```
```{r Heterozygosity}
HET = read.table("~/BTSync/FetalRNAseq/Genome-wide\ genotyping/PLINK_040517_1054/FB_new_with_sex_het.het", h=T, as.is=T)
HET <- HET %>% mutate(H=(N.NM. - O.HOM.) / N.NM.)
ggplot(HET, aes(x=H)) +
  geom_histogram() +
  geom_text(aes(label=IID), y=2, data=filter(HET, H>0.322), angle=-45, vjust=0, hjust=1)

ggplot(HET, aes(x=F)) + 
  geom_histogram() +
  geom_text(aes(label=IID), y=2, data=filter(HET, F< -0.01), angle=45, vjust=0, hjust=0)

filter(HET, F< -.01) %>% left_join(sample_info)
```

# Hardy-Weinberg Equilibrium
- very few loci violating HWE. This might be down to small sample size

```{bash plink_hwe}
cd /Users/heo3/BTSync/FetalRNAseq/Genome-wide\ genotyping/PLINK_040517_1054
~/bin/plink --bfile FB_new_with_sex --hardy  --out FB_new_with_sex_hwe
```

```{r HWE}
HWE = read.table("~/BTSync/FetalRNAseq/Genome-wide\ genotyping/PLINK_040517_1054/FB_new_with_sex_hwe.hwe", h=T, as.is=T)
ggplot(HWE, aes(x=log10(P))) + 
  geom_histogram() 
filter(HWE, P<0.1) %>%
  ggplot(aes(x=log10(P)))+geom_histogram()
```

# PCA

```{bash plinkPCA}
cd /Users/heo3/BTSync/FetalRNAseq/Genome-wide\ genotyping/PLINK_040517_1054
~/bin/plink --bfile FB_new_with_sex --pca  --out FB_new_with_sex_pca
```

``` {r PCA}
PCA = read.table("~/BTSync/FetalRNAseq/Genome-wide\ genotyping/PLINK_040517_1054/FB_new_with_sex_pca.eigenvec", h=F, as.is=T)

ggplot(PCA, aes(x=V3, y=V4))+geom_point()+xlab('PCA1')+ylab('PCA2')
```
