---
title: "eQTL / SMR analysis"
output:
  html_document:
    df_print: paged
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.asp = 0.618, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r setup}
library(tidyverse)
library(ggbeeswarm)
source("plot_SMR.r")
source("FormatGGplot.R")

PlotEQTL<-function(geneID, snp, pos, chr, counts, target, snp_header) {
  snp_pos <- paste(chr, pos, sep=":")
  gene_name=filter(counts, Id==geneID) %>% `$`(SYMBOL)
  expression <- counts %>% filter(Id == geneID) %>%  
    dplyr::select(-Id, -SYMBOL) %>%
    gather("Sample", "value") %>%
    dplyr::select(Sample, value) %>%
    left_join(target)
  genotypes <- system(paste("bcftools view -H -r", snp_pos, "../Shiny/GENEX-FB2/Data/combined_filtered.vcf.gz"), intern=TRUE) %>%
    str_split('\t')
  ref <- genotypes[[1]][4]
  alt <- genotypes[[1]][5]
  data <- tibble(Sample=snp_header[[1]][10:length(genotypes[[1]])], geno=genotypes[[1]][10:length(genotypes[[1]])]) %>%
    separate(geno, c("GT", "DS", "GP" ), sep=":") %>%
    full_join(expression) %>%
    filter(!is.na(GT)) %>%
    mutate(DS=as.numeric(DS), 
           geno=factor(ifelse(GT=='0|0', paste0(UQ(ref), UQ(ref)),
                              ifelse(GT=='1|1', paste0(UQ(alt), UQ(alt)),
                                     paste0(UQ(ref), UQ(alt)))),
                       levels=c(paste0(UQ(ref), UQ(ref)), paste0(UQ(ref), UQ(alt)), 
                                paste0(UQ(alt), UQ(alt)) ))
           )
  #mean <- data %>% group_by(genotype) %>% summarise(mean = mean(value))
  title<-paste0(geneID, ' (', gene_name, ') x ', snp)
  plot<-  ggplot(data, aes(y=value, x=DS)) + 
    #geom_errorbar(aes(ymin=mean, ymax=mean), colour='black', size=1, width=.5, data=mean) +
    geom_smooth(method='lm', colour='black') +
    geom_quasirandom(aes(colour=geno), width=0.05, alpha=.5) + 
    ylab("Normalised Counts") +
    xlab('Dosage') +
    #scale_x_discrete(breaks=c(0,1,2), labels=c('Ref', 'Het', 'Alt')) +
    side_theme() +
    scale_colour_brewer(type = "qual", palette = 6) +
    ggtitle(title) +
    theme(legend.position="right") +
    labs(color='Best Guess\nGenotype')
  plot
}

counts <- read_tsv("../Shiny/GENEX-FB2/Data/counts.txt") %>% select(-Chr, -gene_type, -ChrType)
colnames(counts) <- str_replace_all(colnames(counts), 'norm.', '')
target <- read_tsv('../Data/SampleInfo.txt', col_types = cols(Sample='c'))
snp_header <- system("bcftools view -h ../Shiny/GENEX-FB2/Data/combined_filtered.vcf.gz | tail -1", intern=TRUE) %>%
  str_split('\t')
tx2gene <- read_tsv("~/BTSync/FetalRNAseq/Github/GENEX-FB2/Data/tx2gene.txt",
                    trim_ws = TRUE)
Hauberg <- read_tsv("~/BTSync/FetalRNAseq/Github/GENEX-FB2/Data/Hauberg.txt")
# list of non-schizophrenia/education GWAS tested by Hauberg
other <- c('BMI, hip adjusted', 'BMI, waist adjusted', 'Menopause, age at', 'Height', 'Smoking, cigarettes/day') 
```

# Gene-level analysis

## FastQTL


```{r FastQTL}
egenes <- read_tsv("~/BTSync/FetalRNAseq/Github/GENEX-FB2/FastQTL/egenes.bed.gz",
                   col_names=c("chr", "start", "end", "gene_id", "num_var", "beta_shape1", "beta_shape2", "true_df", "pval_true_df", "variant_id", "tss_distance", "minor_allele_samples", "minor_allele_count", "maf", "ref_factor", "pval_nominal", "slope", "slope_se", "pval_perm", "pval_beta", "qval", "pval_nominal_threshold"))


```

- `r filter(egenes, qval<.1) %>% nrow()` eGenes
- `r format(filter(egenes, qval<.1) %>% nrow/nrow(egenes)*100, digits=2)`% of `r nrow(egenes)` tested genes (4-9% per chromosome)

```{r summarise_FastQTL, dependson='FastQTL'}
filter(egenes, qval<.1) %>% 
  count(chr) %>% 
  rename(eGenes=n) %>% 
  full_join(count(egenes, chr)) %>%
  rename(allGenes=n) %>%
  mutate(proportion=eGenes/allGenes) %>% 
  arrange(desc(proportion))
```


## SMR
- 7 loci have significnat SMR p-values when correcting for the number of loci in the output and p_HEIDI>.05
    - only 806 genes are tested because the p-value cutoff for eQTLs is 5e-8
    - 3 survive when correcting for 29720 tested genes (p<1.68e-06)
    - 3 located in one (dodgy) region of chr15
- 9 additional loci have significant SMR p-values but significant heterogeneity (p_HEIDI<.05)
    - all of these are on Chr6

```{r smr_summary_gene, cache=TRUE}

mysmr <- read_delim("~/BTSync/FetalRNAseq/Github/GENEX-FB2/SMR/mysmr.smr",
"\t", escape_double = FALSE, trim_ws = TRUE)

smr_filter <- select(counts, Id, SYMBOL) %>% right_join(mysmr, by=c('Id' = 'Gene')) %>%
  filter(p_SMR<.05/806 & p_HEIDI>.05) %>% 
  arrange(p_SMR)
smr_filter %>% select(-probeID, -ProbeChr, -Probe_bp, -A1, -A2, -Freq, -b_GWAS, -se_GWAS, -b_eQTL, -se_eQTL, -b_SMR, -se_SMR, -nsnp_HEIDI, -topSNP_bp)

```

## GTex/Common Mind SMR
- HIST1H4L and WBP2NL have not been previously identified by SMR
- GTex eQTLs from all other genes have been linked to Schizophrenia by SMR, but not eQTLs from brain tissue (presumably because brain datasets are small and have fewer eQTLs)
- CSPG4P11 has been identified in SMR using CMC data

```{r gtex_gene, dependson='smr_summary_gene'}
hauberg_counts <- smr_filter %>% left_join(select(Hauberg, GeneID, Trait, eQTL_study), by=c('Id' = 'GeneID')) %>% filter(!is.na(Trait)) %>% count(SYMBOL) %>% rename(total=n)

scz_counts <- filter(Hauberg, Trait=='Schizophrenia') %>%
  select(GeneID, eQTL_study) %>%
  right_join(smr_filter, by=c('GeneID' = 'Id')) %>% 
  mutate(eQTL_study = ifelse(str_detect(eQTL_study, 'Brain'), paste('Scz x', eQTL_study), 'Scz x Other')) %>%
  count(GeneID, eQTL_study) %>% spread(eQTL_study, n) %>%
  select(-`<NA>`)

results <- smr_filter %>% left_join(select(Hauberg, GeneID, Trait, eQTL_study), by=c('Id' = 'GeneID')) %>%
  count(SYMBOL, Trait, Id) %>% 
  spread(Trait, n)

other_counts <- rowSums(select(results, one_of(other)), na.rm=TRUE)
other_counts <- ifelse(other_counts == 0, NA, other_counts)
mutate(results, Other=other_counts) %>%
  select(-one_of(other), -`<NA>`) %>%
  left_join(scz_counts, by=c('Id' = 'GeneID')) %>%
  select(SYMBOL, starts_with('Scz'), everything(), -Schizophrenia, -Id)
```

### HIST1H4L (histone cluster 1 H4 family member l)
- CLOZUK locus 1 (xMHC)
```{r HIST1H4L, cache=TRUE}
#smr --bfile ../Genotypes/Plink/genotypes --gwas-summary CLOZUK_recoded.txt --beqtl-summary mybesd --out HIST1H4L  --plot --probe ENSG00000275126 --probe-wind 500 --gene-list ../Data/geneloc_smr.txt
SMRData = ReadSMRData("../SMR/HIST1H4L.ENSG00000275126.txt")

SMRData[["SMR"]][["V4"]] <- counts[sapply(SMRData[["SMR"]][["V4"]], function(x) which(counts$Id ==x)),]$SYMBOL

SMRLocusPlot(data=SMRData, smr_thresh=.05/806, heidi_thresh=0.05, plotWindow=500, max_anno_probe=16)
SMRLocusPlot(data=SMRData, smr_thresh=.05/806, heidi_thresh=0.05, plotWindow=10, max_anno_probe=16)
SMREffectPlot(data=SMRData, trait_name="CLOZUK")
PlotEQTL('ENSG00000275126', 'rs200995', 27845916, 6, counts, target, snp_header)

```

## AC243562.2 (processed pseudogene) / CSPG4P11 (Chondroitin Sulfate Proteoglycan 4 Pseudogene 11) / UBE2Q2L (ubiquitin conjugating enzyme E2 Q2 like)
- CLOZUK 76 

```{r UBE2Q2L, cache=TRUE}
#smr --bfile ../Genotypes/Plink/genotypes --gwas-summary CLOZUK_recoded.txt --beqtl-summary mybesd --out AC243562  --plot --probe ENSG00000259683 --probe-wind 500 --gene-list ../Data/geneloc_smr.txt
SMRData = ReadSMRData("../SMR/AC243562.ENSG00000259683.txt")
SMRLocusPlot(data=SMRData, smr_thresh=.05/806, heidi_thresh=0.05, plotWindow=500, max_anno_probe=16)
SMRLocusPlot(data=SMRData, smr_thresh=.05/806, heidi_thresh=0.05, plotWindow=250, max_anno_probe=16)
SMREffectPlot(data=SMRData, trait_name="sczx")
PlotEQTL('ENSG00000259683', 'rs4502182', 84254768, 15, counts, target, snp_header)
PlotEQTL('ENSG00000259726', 'rs6603022', 84161418, 15, counts, target, snp_header)
PlotEQTL('ENSG00000259511', 'rs62028142', 84386630, 15, counts, target, snp_header)
```


## LINC01068
- CLOZUK 105 (NDFIP2, NDFIP2-AS1, RBM26, RBM26-AS1)
```{r LINC01068, cache=TRUE}
#smr --bfile ../Genotypes/Plink/genotypes --gwas-summary CLOZUK_recoded.txt --beqtl-summary mybesd --out LINC01068  --plot --probe ENSG00000227676 --probe-wind 500 --gene-list ../Data/geneloc_smr.txt
SMRData = ReadSMRData("../SMR/LINC01068.ENSG00000227676.txt")
SMRLocusPlot(data=SMRData, smr_thresh=.05/806, heidi_thresh=0.05, plotWindow=500, max_anno_probe=16)
SMRLocusPlot(data=SMRData, smr_thresh=.05/806, heidi_thresh=0.05, plotWindow=25, max_anno_probe=16)
SMREffectPlot(data=SMRData, trait_name="sczx") 
PlotEQTL('ENSG00000227676', 'rs1212082', 79552633, 13, counts, target, snp_header)
```

## ABCB9 (ATP binding cassette subfamily B member 9)
- CLOZUK 11
```{r ABCB9, cache=TRUE}
#smr --bfile ../Genotypes/Plink/genotypes --gwas-summary CLOZUK_recoded.txt --beqtl-summary mybesd --out ABCB9  --plot --probe ENSG00000150967 --probe-wind 500 --gene-list ../Data/geneloc_smr.txt
SMRData = ReadSMRData("../SMR/ABCB9.ENSG00000150967.txt")
SMRLocusPlot(data=SMRData, smr_thresh=.05/806, heidi_thresh=0.05, plotWindow=500, max_anno_probe=16)
SMRLocusPlot(data=SMRData, smr_thresh=.05/806, heidi_thresh=0.05, plotWindow=100, max_anno_probe=16)
SMREffectPlot(data=SMRData, trait_name="sczx")
PlotEQTL('ENSG00000150967', 'rs1790106', 123107049, 12, counts, target, snp_header)
```

## WBP2NL (WBP2 N-terminal like)
- CLOZUK 14
```{r WBP2NL, cache=TRUE}
#smr --bfile ../Genotypes/Plink/genotypes --gwas-summary CLOZUK_recoded.txt --beqtl-summary mybesd --out WBP2NL  --plot --probe ENSG00000183066 --probe-wind 500 --gene-list ../Data/geneloc_smr.txt
SMRData = ReadSMRData("../SMR/WBP2NL.ENSG00000183066.txt")
SMRLocusPlot(data=SMRData, smr_thresh=.05/806, heidi_thresh=0.05, plotWindow=500, max_anno_probe=16)
SMRLocusPlot(data=SMRData, smr_thresh=.05/806, heidi_thresh=0.05, plotWindow=100, max_anno_probe=16)
SMREffectPlot(data=SMRData, trait_name="sczx")
PlotEQTL('ENSG00000183066', 'rs5996092', 41948404, 22, counts, target, snp_header)
```

# Transcript-level analysis
## FastQTL

```{r FastQTL_tr, cached=TRUE}
egenes_tr <- read_tsv("~/BTSync/FetalRNAseq/Github/GENEX-FB2/FastQTL/egenes_transcript.bed.gz",
                      col_names=c("chr", "start", "end", "gene_id", "num_var", "beta_shape1", "beta_shape2", "true_df", "pval_true_df", "variant_id", "tss_distance", "minor_allele_samples", "minor_allele_count", "maf", "ref_factor", "pval_nominal", "slope", "slope_se", "pval_perm", "pval_beta", "qval", "pval_nominal_threshold"))

egenes_tr <- egenes_tr %>% rename(transcript_id=gene_id) %>% left_join(tx2gene) 
```

- `r filter(egenes_tr, qval<.1) %>% nrow()` eTranscripts belonging to `r filter(egenes_tr, qval<.1) %>% distinct(gene_id) %>% nrow()` distinct genes
- `r format(filter(egenes_tr, qval<.1) %>% nrow/nrow(egenes_tr)*100, digits=2)`% of `r nrow(egenes_tr)` tested transcripts (2.2% - 3.6% per chromosome)

```{r summarise_FastQTL_tr, dependson='FastQTL_tr'}
filter(egenes_tr, qval<.1) %>% summarise(n=n()/nrow(egenes_tr)) %>% `$`(n)


filter(egenes_tr, qval<.1) %>% 
  count(chr) %>% 
  rename(eGenes=n) %>% 
  full_join(count(egenes_tr, chr)) %>%
  rename(allGenes=n) %>%
  mutate(proportion=eGenes/allGenes) %>% 
  arrange(desc(proportion))


```

## SMR
- 13 transcripts from 12 genes have significnat SMR p-values when correcting for the number of transcripts in the output and p_HEIDI>.05
- all but two of loci that were significant in the gene-level analysis remain significant, though the top SNP is different in some cases and the p-values are slightly different
    - WBP2NL and UBE2Q2L are non-significant because of the more stringent bonforoni correction (2042 tests vs. 806)
- 3 loci in xMHC
- 3 loci in CLOZUK 76 (chr15), plus another locus 6 mb away
- additional loci on chr 14 (top hit), chr3 and chr11
    
```{r smr_summary_tr, cache=TRUE}
mysmr_tr <- read_tsv("~/BTSync/FetalRNAseq/Github/GENEX-FB2/SMR/mysmr_transcript_all.smr.gz",
                     col_names = c("probeID", "ProbeChr", "Gene", "Probe_bp", "topSNP",
                                   "topSNP_chr", "topSNP_bp", "A1", "A2", "Freq", "b_GWAS",
                                   "se_GWAS", "p_GWAS", "b_eQTL", "se_eQTL",
                                   "p_eQTL", "b_SMR", "se_SMR", "p_SMR", "p_HEIDI",
                                   "nsnp_HEIDI"),
                     trim_ws = TRUE)

smr_tr <- mysmr_tr %>% mutate(topSNP_mb=topSNP_bp/1000000) %>% 
  select(-probeID, -ProbeChr, -Probe_bp, -A1, -A2, -Freq, -b_GWAS, -se_GWAS, -b_eQTL, -se_eQTL, -b_SMR, -se_SMR, -nsnp_HEIDI, -topSNP_bp) %>%
  rename(transcript_id = Gene)

smr_filt_tr <- smr_tr %>% filter(p_SMR < 0.05/nrow(smr_tr) & p_HEIDI>.05)  %>%
  left_join(tx2gene) %>%
  left_join(select(counts, Id, SYMBOL), by=c('gene_id' = 'Id'))

smr_filt_tr %>%  group_by(SYMBOL) %>%
  slice(1) %>%
  ungroup() %>%
  left_join(count(smr_filt_tr, gene_id)) %>%
  select(SYMBOL, num_tr=n, topSNP, topSNP_mb, everything(), -transcript_id, -gene_id) %>%
  arrange(p_SMR)

```

## GTex/Common Mind SMR
- FLOT1, MSANTD2 and RNF5 have not been previously identified by SMR
- GTex eQTLs from 4 additional genes not found in the gene-level analysis have been linked to Schizophrenia by SMR (GNL3, KLC1, PSMA4 and RN7SL417P)
    - not eQTLs from brain tissue (presumably because brain datasets are small and have fewer eQTLs)



```{r gtex_tr, dependson='smr_summary_tr'}
hauberg_counts_tr <- smr_filt_tr %>% left_join(select(Hauberg, GeneID, Trait, eQTL_study), by=c('gene_id' = 'GeneID')) %>% filter(!is.na(Trait)) %>% count(SYMBOL) %>% rename(total=n)

scz_counts_tr <- filter(Hauberg, Trait=='Schizophrenia') %>%
  select(GeneID, eQTL_study) %>%
  right_join(smr_filt_tr, by=c('GeneID' = 'gene_id')) %>%
  mutate(eQTL_study = ifelse(str_detect(eQTL_study, 'Brain'), paste('Scz x', eQTL_study), 'Scz x Other')) %>%
  count(GeneID, eQTL_study) %>% spread(eQTL_study, n) %>%
  select(-`<NA>`)

results_tr <- smr_filt_tr %>% left_join(select(Hauberg, GeneID, Trait, eQTL_study), by=c('gene_id' = 'GeneID')) %>%
  count(SYMBOL, Trait, gene_id) %>% 
  spread(Trait, n)

other_counts_tr <- rowSums(select(results_tr, one_of(other)), na.rm=TRUE)
other_counts_tr <- ifelse(other_counts_tr == 0, NA, other_counts_tr)
mutate(results_tr, Other=other_counts_tr) %>%
  select(-one_of(other), -`<NA>`) %>%
  left_join(scz_counts_tr, by=c('gene_id' = 'GeneID')) %>%
  select(SYMBOL, starts_with('Scz'), everything(), -Schizophrenia, -gene_id)
```

## KLC1
- CLOZUK 17

